---
title: "Axis Customization"
author: "Abdullah Al Shamim"
format: html
editor: visual
---

## Axis Customization

#### Data Preparation

```{r}

# Reproducible data
set.seed(1)
x <- runif(100, min = -5, max = 5)
y <- x ^ 3 + rnorm(100, mean = 0, sd = 5)

```

## X and Y axis labels 

The **default axis labels will depend on the function you are using**, e.g. `plot` function will use the names of the input data, `boxplot` won’t show any axis labels by default and `hist` will show the name of the variable on the X-axis and “Frequency” or “Density” on the Y-axis, depending on the type of the histogram.

#### Default plot

```{r message=FALSE, error=FALSE, warning=FALSE}


plot(x, y, pch = 25)
```

#### Custom axes labels

In order to change the axis labels you can specify the arguments `xlab` and `ylab` as follows:

```{r message=FALSE, error=FALSE, warning=FALSE}

plot(x, y, pch = 19,
     xlab = "My X-axis label",
     ylab = "My Y-axis label") 
 
```

#### Labels color

In addition, is is possible to modify the color of the axes labels with the `col.lab` argument.

```{r message=FALSE, error=FALSE, warning=FALSE}

plot(x, y, pch = 19,
     xlab = "My X-axis label",
     ylab = "My Y-axis label",
     col.lab = 4) # Color of the labels 

```

###  Remove axis labels 

You can remove the axis labels with two different methods:

#### **Option 1**. Set the `xlab` and `ylab` arguments to `""`, `NA` or `NULL`.

```{r message=FALSE, error=FALSE, warning=FALSE}

# Delete labels
plot(x, y, pch = 19,
     xlab = "", # Also NA or NULL
     ylab = "") # Also NA or NULL 
```

#### **Option 2**. Set the argument `ann` to `FALSE`. This will override the label names if provided.

```{r message=FALSE, error=FALSE, warning=FALSE}

# Equivalent
plot(x, y, pch = 19,
     xlab = "My X-axis label",
     ylab = "My Y-axis label",
     ann = FALSE) 

```

## The `axis` function 

The `axis` function **allows adding axes to all sides
of the the current plot, with the possibility of specifying the position
and the labels of the tick marks**. Recall to type `?axis` for further details.

![](images/Screenshot (1233).png){width="344"}

**Option 1**. Set `xaxt = "n"` and `yaxt = "n"` to remove the tick labels of the plot and add the new labels with the `axis` function. Note that the `at` argument sets where to show the tick marks.

```{r message=FALSE, error=FALSE, warning=FALSE}

plot(x, y, pch = 19,
     xaxt = "n", yaxt = "n")

# X-axis
axis(1, at = c(-4, 0, 4))

# Y-axis
axis(2, at = c(-100, -50, 0, 50, 100)) 

```

**Option 2**. Set `axes = FALSE` inside your plotting function to remove the plot box and add the new axes with the `axis` function.

```{r message=FALSE, error=FALSE, warning=FALSE}

plot(x, y, pch = 19,
     axes = FALSE)

# Add X-axis
axis(1)

# Add Y-axis
axis(2)

# You can add the plot box again
# box() 

```

#### Color of the axes

You can also customize the color of the axis and the ticks with the `col` and `col.ticks` arguments.

```{r message=FALSE, error=FALSE, warning=FALSE}

plot(x, y, pch = 19,
     axes = FALSE)

# Add X-axis
axis(1,
     col = "blue",        # Axis line color
     col.ticks = "green", # Ticks color
     col.axis = "red")    # Labels color

# Add Y-axis
axis(2,
     col = "blue",
     col.ticks = "green",
     col.axis = "red") 

```

## Axis tick marks

In addition to axes labels, the tick marks of each axis can be customized in different ways.

### `xaxp` and `yaxp` arguments 

The `xaxp` (X-axis) and `yaxp` (Y-axis) arguments
allow customizing where the ticks of each axis start and end and the
number of regions to divide the axis specifying vectors of the form `c(start, end, number_regions)`.

```{r message=FALSE, error=FALSE, warning=FALSE}

plot(x, y, pch = 19,
     xaxp = c(-3, 3, 3),
     yaxp = c(-70, 70, 5)) 
 
```

### Minor ticks

It is possible to add minor ticks to the axes with the `minor.tick` function of the `Hmisc` library. The function will allow you to specify the tick density, the size and addition arguments to each axis.

```{r message=FALSE, error=FALSE, warning=FALSE}

# install.packages("Hmisc")
library(Hmisc)

plot(x, y, pch = 19)
minor.tick(nx = 2, ny = 2,   # Ticks density
           tick.ratio = 0.5) # Ticks size 
 

```

### Interior ticks

You can set interior ticks passing positive values to `tck` argument. The greater the value, the longer the ticks. Default value is `tck = -0.5`.

```{r message=FALSE, error=FALSE, warning=FALSE}

# Interior ticks
plot(x, y, pch = 19,
     tck = 0.02) 

```

### Rotate tick mark labels

It is possible to rotate the tick mark labels in several ways making use of the `las` argument.

#### **Option 1**. Parallel to axis (default).

এখন এই Parallel-কে Horizontal, Perpendicular, Vertical–করতে main = "Parallel"-এর পরিবর্তে তা বসিয়ে দিন।

```{r message=FALSE, error=FALSE, warning=FALSE}

plot(x, y, pch = 19,
     las = 0,
     main = "Parallel") 

```

#### **Option 2**. Horizontal.

```{r}

plot(x, y, pch = 19,
     las = 1,
     main = "Horizontal") 

```

#### **Option 3**. Perpendicular to axis

```{r}

plot(x, y, pch = 19,
     las = 2,
     main = "Perpendicular") 

```

#### **Option 4**. Vertical.

```{r}

plot(x, y, pch = 19,
     las = 3,
     main = "Vertical") 

```

### Custom tick mark labels 

The `labels` argument of the `axis` function allows customizing the tick mark labels.

```{r message=FALSE, error=FALSE, warning=FALSE}

# Change X-axis tick labels
plot(x, y, pch = 19, xaxt = "n")
axis(1, at = seq(round(min(x)),
                 round(max(x)), by = 1),
     labels = 1:11) 

```

Note that you can also set texts in addition to numbers, or even expressions.

```{r message=FALSE, error=FALSE, warning=FALSE}

# X-axis tick labels with text
plot(x, y, pch = 19, xaxt = "n")
axis(1, at = c(-2, 2),
     labels = c("Some text", "Other text")) 

```

### Remove tick marks

#### **Option 1**. Remove only X-axis ticks.

```{r message=FALSE, error=FALSE, warning=FALSE}

# Remove X axis tick labels
plot(x, y, pch = 19,
     xaxt = "n",
     main = "xaxt = 'n'") 

```

#### **Option 2**. Remove only Y-axis ticks.

```{r message=FALSE, error=FALSE, warning=FALSE}

# Remove Y axis tick labels
plot(x, y, pch = 19,
     yaxt = "n",
     main = "yaxt = 'n'") 

```

#### Option 3. Remove both X-axis and Y-axis ticks.

```{r message=FALSE, error=FALSE, warning=FALSE}

# Remove both axis tick labels
plot(x, y, pch = 19,
     yaxt = "n",
     xaxt = "n",
     main = "xaxt = 'n', yaxt = 'n'")
```

#### **Option 4**. Remove both X-axis and Y-axis ticks and the box.

```{r message=FALSE, error=FALSE, warning=FALSE}

# Remove axis tick labels and the box
plot(x, y, pch = 19,
     axes = FALSE,
     main = "axes = FALSE") 
 

```

## Axis limits 

The axis limits can be customized with the `xlim` and `ylim` arguments for the X and Y axis, respectively.

```{r message=FALSE, error=FALSE, warning=FALSE}

# Axis limits
plot(x, y, pch = 19,
     xlim = c(0, 4),   # X-axis limits
     ylim = c(0, 100)) # Y-axis limits 
 
```

## Dual axis  

You can join together two plots with different Y-axis scale increasing the margins of the plot, using `par(new = TRUE)`, creating a new plot without axis or labels and setting a new axis with the `axis` function.

```{r message=FALSE, error=FALSE, warning=FALSE}

# Increase the plot margins
par(mar = c(5, 4, 4, 4) + 0.25)

# Data
plot(x, y, pch = 19, ylab = "Var 1")

# Needed to merge the plots
par(new = TRUE)

# More data
plot(x ^ 2, y ^ 2,
     col = 4, pch = 19,
     axes = FALSE, # No axes
     bty = "n",    # No box
     xlab = "", ylab = "") # No axis labels

# New axis
axis(4)

# Axis label
mtext("Var 2", side = 4, line = 3, col = 4) 

```

```{r message=FALSE, error=FALSE, warning=FALSE}

```

```{r message=FALSE, error=FALSE, warning=FALSE}

```
