---
title: "Histogram Basics"
author: "Abdullah Al Shamim"
format: html
editor: visual
---

## Histogram

```{r message= FALSE, error=FALSE, warning=FALSE}

library(tidyverse)

msleep %>%
  ggplot(aes(x = awake)) +
  geom_histogram() + 
  labs(title = "Histogram of Total Sleep",
       x = "Total Sleep",
       y = "NULL") +
  theme_bw()
```

You can add options to executable code like this

```{r message= FALSE, error=FALSE, warning=FALSE}

msleep %>%
  ggplot(aes(x = awake)) +
  geom_histogram(binwidth = 2, fill = "#97B3C6") + 
  labs(title = "Histogram of Total Sleep",
       x = "Total Sleep",
       y = "NULL") +
  theme_bw()

```

#### A basic histogram

```{r message= FALSE, error=FALSE, warning=FALSE}

# Sample data
set.seed(3)
x <- rnorm(200)

# Histogram
hist(x, prob = TRUE)
```

#### Histogram with normal curve 

```{r message= FALSE, error=FALSE, warning=FALSE}

# X-axis grid
x2 <- seq(min(x), max(x), length = 40)

# Normal curve
fun <- dnorm(x2, mean = mean(x), sd = sd(x))

# Histogram
hist(x, prob = TRUE, col = "white",
     ylim = c(0, max(fun)),
     main = "Histogram with normal curve")
lines(x2, fun, col = 2, lwd = 2) 
 
```

#### Histogram with density line 

```{r message= FALSE, error=FALSE, warning=FALSE}

# Sample data
set.seed(3)
x <- rnorm(200)

# Normal curve
fun <- dnorm(x2, mean = mean(x), sd = sd(x))

# Histogram
hist(x, prob = TRUE, ylim = c(0, max(fun)),
     main = "Histogram with density curve")
lines(density(x), col = 4, lwd = 2) #lwd = line width

```

#### Histogram with mean line 

```{r message= FALSE, error=FALSE, warning=FALSE}

#install.packages("ggpubr")
library(ggpubr)


# Create some data format
set.seed(1234)
men = data.frame(
  sex = factor(rep(c("F", "M"), each=200)),
  weight = c(rnorm(200, 55), rnorm(200, 58)))
head(men, 4)

# Use custom color palette
# Change outline and fill colors by groups ("sex")
gghistogram(men, x = "weight",
            add = "mean", #rug = TRUE,
            color = "sex", fill = "sex",
            palette = c("#00AFBB", "#E7B800"))


```

### Data preparation

```{r message= FALSE, error=FALSE, warning=FALSE}

set.seed(3)
x1 <- rnorm(500)
x2 <- rnorm(500, mean = 3)
x <- c(x1, x2)
group <- c(rep("G1", 500), rep("G2", 500))

df <- data.frame(x, group = group)

view(df)
```

#### a) Grouped histogram with `geom_histogram`

```{r message= FALSE, error=FALSE, warning=FALSE}

# install.packages("ggplot2")
library(ggplot2)

# Histogram by group in ggplot2
ggplot(df, aes(x = x, fill = group)) + 
  geom_histogram()

```

#### b) Colour

```{r message= FALSE, error=FALSE, warning=FALSE}


# install.packages("ggplot2")
library(ggplot2)

# Histogram by group in ggplot2
ggplot(df, aes(x = x, colour = group)) + 
  geom_histogram()
```

#### c) identity position

```{r message= FALSE, error=FALSE, warning=FALSE}

# install.packages("ggplot2")
library(ggplot2)

# Histogram by group in ggplot2
ggplot(df, aes(x = x, fill = group, colour = group)) + 
  geom_histogram(alpha = 0.5, position = "identity") 


```

#### d) dodge position

```{r message= FALSE, error=FALSE, warning=FALSE}

# install.packages("ggplot2")
library(ggplot2)

# Histogram by group in ggplot2
ggplot(df, aes(x = x, fill = group, colour = group)) + 
  geom_histogram(position = "dodge") 


```

#### e) Histogram by group with custom colors

```{r message= FALSE, error=FALSE, warning=FALSE}

# install.packages("ggplot2")
library(ggplot2)

# Histogram by group in ggplot2
ggplot(df, aes(x = x, fill = group)) + 
  geom_histogram(colour = "black",
                 lwd = 0.75,
                 linetype = 1,
                 position = "identity") 
 

```

#### f) Fill color

```{r message= FALSE, error=FALSE, warning=FALSE}

# install.packages("ggplot2")
library(ggplot2)

# Histogram by group in ggplot2
ggplot(df, aes(x = x, colour = group)) + 
  geom_histogram(fill  = "white",
                 position = "identity") 

```

#### g) Custom border colors for each group

```{r message= FALSE, error=FALSE, warning=FALSE}

# install.packages("ggplot2")
library(ggplot2)

# Histogram by group in ggplot2
ggplot(df, aes(x = x, colour = group)) + 
  geom_histogram(fill  = "white",
                 position = "identity") +
  scale_color_manual(values = c("blue", "orange")) 


```

#### h) Custom fill colors for each group

```{r message= FALSE, error=FALSE, warning=FALSE}

# install.packages("ggplot2")
library(ggplot2)

# Histogram by group in ggplot2
ggplot(df, aes(x = x, fill = group)) + 
  geom_histogram(color = 1, alpha = 0.75,
                 position = "identity") +
  scale_fill_manual(values = c("#8795E8", "#FF6AD5")) 

```

#### i) Legend customization

Custom legend title

```{r message= FALSE, error=FALSE, warning=FALSE}

# install.packages("ggplot2")
library(ggplot2)

# Histogram by group in ggplot2
ggplot(df, aes(x = x, fill = group, colour = group)) + 
  geom_histogram(alpha = 0.5, position = "identity") + 
  guides(fill = guide_legend(title = "Title"),
         colour = guide_legend(title = "Title")) 


```

Custom legend labels

```{r message= FALSE, error=FALSE, warning=FALSE}

# install.packages("ggplot2")
library(ggplot2)

# Histogram by group in ggplot2
ggplot(df, aes(x = x, fill = group, colour = group)) + 
  geom_histogram(alpha = 0.5, position = "identity") + 
  scale_color_discrete(labels = c("A", "B")) +
  scale_fill_discrete(labels = c("A", "B")) 

```

Legend position

```{r message= FALSE, error=FALSE, warning=FALSE}

# install.packages("ggplot2")
library(ggplot2)

# Histogram by group in ggplot2
ggplot(df, aes(x = x, fill = group, colour = group)) + 
  geom_histogram(alpha = 0.5, position = "identity") + 
  theme(legend.position = "left") 

```

Remove the legend

```{r message= FALSE, error=FALSE, warning=FALSE}

# install.packages("ggplot2")
library(ggplot2)

# Histogram by group in ggplot2
ggplot(df, aes(x = x, fill = group, colour = group)) + 
  geom_histogram(alpha = 0.5, position = "identity") + 
  theme(legend.position = "none") 
```
